Forecasting climate change impacts on plant population dynamics at large spatial extents: a test case with sagebrush (\emph{Artemisia}) species

### 2017-02-27 docs/ms_sageAbundance.tex
just changes font color

\documentclass[12pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
    \usepackage{xltxtra,xunicode}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={},
            colorlinks=true,
            citecolor=blue,
            urlcolor=black,
            linkcolor=black,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{0}

@@

\maketitle

@@

@@

Forecasting the impacts of climate change on plant populations and
communities is a central challenge for ecology (Clark et al. 2001,
Petchey et al. 2015). Population models are ideally suited for meeting
such a challenge because they provide a way to link climate drivers
directly to population dynamics (Hare et al. 2010, Adler et al. 2012,
Ross et al. 2015, Shriver 2015). However, inference from population
models is typically limited to small spatial extents because the data
required is difficult to collect across broad species ranges. Almost
every study of plant population dynamics relies on demographic
observations recorded at the meter to sub-meter scale (see, e.g.,
Salguero-G{ó}mez et al. 2015). Local-scale demographic data make
building population projection models an easy task (Ellner and Rees
2006, Rees and Ellner 2009, Adler et al. 2012), but it is very difficult
to extrapolate small-scale studies to large spatial extents with any
certainty because the data likely only represent a small subset of
parameter space and environmental conditions (Freckleton et al. 2011,
Queenborough et al. 2011). The real challenge is not to simply make
population forecasts, but to do so at spatial scales relevant to policy
and management decisions (Queenborough et al. 2011).

@@

Large-scale, spatially-explicit population models based on RS data could
offer a valuable new way to investigate the effects of large-scale
environmental changes playing out at landscape and regional scales. Most
current assesments of how plant and animal populations will respond to
climate change rely on species distribution models (SDMs). SDMs rely on
static associations between contemporary climate and a species'
distribution or, more rarely, abundance to project future distribution
or abundance (Elith and Leathwick 2009) and they are easily applied at
landscape to continental scales (e.g., Maiorano et al. 2013, Clark et
al. 2014). However, the short-term and small-scale population dynamics
that actually drive the large-scale distributions of species are not
represented in most SDMs. Because SDMs typically rely on occurence data,
their projections of habitat suitability or probability of occurence
provide little information on the future states of populations in the
core of their range -- areas where a species exists now and is expected
to persist in the future (Ehrl{é}n and Morris 2015). Furthermore,
because they lack short-term dynamics, SDMs usually cannot produce any
estimate of the rate at which local populations will increase or
decrease in the near-term and instead project a future equilibrium
species distribution that may or may not ever be reached. Direct
validation of such predictions is extremely rare (Roberts and Hamann
2012). Large-scale dynamic population models could overcome these
limitations. They would produce spatially-explicit estimates of species
abundance within the species range (Ehrl{é}n and Morris 2015), have the
potential to model expansion in abundance outside the range when coupled
with dynamic models of dispersal, and would provide testable predictions
of how populations should respond to short-term climate perturbations.
These short-term predictions also would give modelers the opportunity to
repeatedly validate and refine their models (Luo et al. 2011).

@@

Adler, P. B., H. J. Dalgleish, and S. P. Ellner. 2012. Forecasting plant
community impacts of climate variability and change: when do competitive
interactions matter? Journal of Ecology 100:478--487.

Apodaca, L. F. 2013. Assessing Growth Response to Climate Controls in a
Great Basin Artemisia Tridentata Plant Community. PhD thesis, University
of Nevada Las Vegas.

Arnett, E. B., and T. Z. Riley. 2015. Science, policy, and the fate of
the greater sage-grouse. Frontiers in Ecology and the Environment
13:235.

Baldeck, C. A., and G. P. Asner. 2014. Improving remote species
identification through efficient training data collection. Remote
Sensing 6:2682--2698.

Barry, R. P., and J. M. V. Hoef. 1996. Blackbox Kriging: Spatial
Prediction without Specifying Variogram Models.

Bradford, J. B., and W. K. Lauenroth. 2006. Controls over invasion of
Bromus tectorum: The importance of climate, soil, disturbance, and seed
availability. Journal of Vegetation Science 17:693--704.

Bradley, B. A. 2010. Assessing ecosystem threats from global and
regional change: hierarchical modeling of risk to sagebrush ecosystems
from climate change, land use and invasive species in Nevada, USA.
Ecography 33:198--208.

Bradshaw, G. A., and J. G. Borchers. 2000. Uncertainty as information:
Narrowing the science-policy gap. Ecology and Society 4.

Clark, J. S., D. M. Bell, M. H. Hersh, and L. Nichols. 2011. Climate
change vulnerability of forest biodiversity: Climate and competition
tracking of demographic rates. Global Change Biology 17:1834--1849.

Clark, J. S., S. R. Carpenter, M. Barber, S. Collins, A. Dobson, J. A.
Foley, D. M. Lodge, M. Pascual, R. Pielke, W. Pizer, C. Pringle, W. V.
Reid, K. A. Rose, O. Sala, W. H. Schlesinger, D. H. Wall, and D. Wear.
2001. Ecological forecasts: an emerging imperative. Science (New York,
N.Y.) 293:657--660.

Clark, J. S., A. E. Gelfand, C. W. Woodall, and K. Zhu. 2014. More than
the sum of the parts: Forest climate response from joint species
distribution models. Ecological Applications 24:990--999.

Colgan, M. S., and G. P. Asner. 2014. Coexistence and environmental
filtering of species-specific biomass in an African savanna. Ecology
95:1579--1590.

Conn, P. B., D. S. Johnson, J. M. V. Hoef, M. B. Hooten, J. M. London,
and P. L. Boveng. 2015. Using spatiotemporal statistical models to
estimate animal abundance and infer ecological dynamics from survey
counts. Ecological Monographs 85:235--252.

Dalgleish, H. J., D. N. Koons, M. B. Hooten, C. A. Moffet, and P. B.
Adler. 2011. Climate influences the demography of three dominant
sagebrush steppe plants. Ecology 92:75--85.

Debinski, D. M., H. Wickham, K. Kindscher, J. C. Caruthers, and M.
Germino. 2010. Montane meadow change during drought varies with
background hydrologic regime and plant functional group. Ecology
91:1672--81.

Ehrl{é}n, J., and W. F. Morris. 2015. Predicting changes in the
distribution and abundance of species under environmental change.
Ecology Letters 18:303--314.

Elith, J., and J. R. Leathwick. 2009. Species Distribution Models:
Ecological Explanation and Prediction Across Space and Time.

Ellner, S. P., and M. Rees. 2006. Integral projection models for species
with complex demography. The American naturalist 167:410--428.

Freckleton, R. P., W. J. Sutherland, A. R. Watkinson, and S. A.
Queenborough. 2011. Density-structured models for plant population
dynamics. American Naturalist 177:1--17.

Gelman, A., and J. Hill. 2009. Data analysis using regression and
multilevel/hierarchical models. Cambridge University Press, Cambridge.

Gelman, A., and D. B. Rubin. 1992. Inference from Iterative Simulation
Using Multiple Sequences. Statistical Science 7:457--472.

Gerber, B. D., W. L. Kendall, M. B. Hooten, J. A. Dubovsky, and R. C.
Drewien. 2015. Optimal population prediction of sandhill crane
recruitment based on climate-mediated habitat limitations. Journal of
Animal Ecology 84:1299--1310.

Germino, M. J., and K. Reinhardt. 2014. Desert shrub responses to
experimental modification of precipitation seasonality and soil depth:
Relationship to the two-layer hypothesis and ecohydrological niche.
Journal of Ecology 102:989--997.

Hare, J. a, M. a Alexander, M. J. Fogarty, E. H. Williams, and J. D.
Scott. 2010. Forecasting the dynamics of a coastal fishery species using
a coupled climate--population model. Ecological applications : a
publication of the Ecological Society of America 20:452--464.

Harte, J., S. R. Saleska, and C. Levy. 2015. Convergent ecosystem
responses to 23-year ambient and manipulated warming link advancing
snowmelt and shrub encroachment to transient and long-term climate-soil
carbon feedback. Global change biology 21:2349--56.

He, K. S., B. A. Bradley, A. F. Cord, D. Rocchini, M.-N. Tuanmu, S.
Schmidtlein, W. Turner, M. Wegmann, and N. Pettorelli. 2015. Will remote
sensing shape the next generation of species distribution models? Remote
Sensing in Ecology and Conservation 1:4--18.

Higdon, D. 1998. A process-convolution approach to modelling
temperatures in the North Atlantic Ocean. Environmental and Ecological
Statistics 5:173--190.

Higdon, D. M. 2002. Space and space-time modeling using process
convolutions. Pages 37--56 \emph{in} C. Anderson, V. Barnett, P.
Chatwin, and A. El-Shaarawi, editors. Quantitative methods for current
environmental issues. Springer, London.

Hobbs, N. T., and M. B. Hooten. 2015. Bayesian Models: A Statistical
Primer for Ecologists. Princeton University Press, Princeton.

Homer, C. G., C. L. Aldridge, D. K. Meyer, and S. J. Schell. 2012.
Multi-scale remote sensing sagebrush characterization with regression
trees over Wyoming, USA: Laying a foundation for monitoring.
International Journal of Applied Earth Observation and Geoinformation
14:233--244.

Homer, C. G., G. Xian, C. L. Aldridge, D. K. Meyer, T. R. Loveland, and
M. S. O'Donnell. 2015. Forecasting sagebrush ecosystem components and
greater sage-grouse habitat for 2050: Learning from past climate
patterns and Landsat imagery to predict the future. Ecological
Indicators 55:131--145.

Hooten, M. B., and N. T. Hobbs. 2015. A guide to Bayesian model
selection for ecologists. Ecological Monographs 85:3--28.

Hooten, M. B., and C. K. Wikle. 2007. Shifts in the spatio-temporal
growth dynamics of shortleaf pine. Environmental and Ecological
Statistics 14:207--227.

Hooten, M. B., D. R. Larsen, and C. K. Wikle. 2003. Predicting the
spatial distribution of ground flora on large domains using a
hierarchical Bayesian model. Landscape Ecology 18:487--502.

Kuchler, A. 1964. Potential Natural Vegetation of the Conterminous
United States. American Geographical Society, Special Publication No.
36.

Latimer, A. M., S. Banerjee, H. Sang, E. S. Mosher, and J. A. Silander.
2009. Hierarchical models facilitate spatial analysis of large data
sets: A case study on invasive plant species in the northeastern United
States. Ecology Letters 12:144--154.

Luo, Y., K. Ogle, C. Tucker, S. Fei, C. Gao, S. LaDeau, J. S. Clark, and
D. S. Schimel. 2011. Ecological forecasting and data assimilation in a
data-rich era. Ecological Applications 21:1429--1442.

Maiorano, L., R. Cheddadi, N. E. Zimmermann, L. Pellissier, B.
Petitpierre, J. Pottier, H. Laborde, B. I. Hurdu, P. B. Pearman, A.
Psomas, J. S. Singarayer, O. Broennimann, P. Vittoz, A. Dubuis, M. E.
Edwards, H. A. Binney, and A. Guisan. 2013. Building the niche through
time: using 13,000 years of data to predict the effects of climate
change on three tree species in Europe. Global Ecology and Biogeography
22:302--317.

Merow, C., A. M. Latimer, A. M. Wilson, S. M. McMahon, A. G. Rebelo, and
J. A. Silander. 2014. On using integral projection models to generate
demographically driven predictions of species' distributions:
development and validation using sparse data. Ecography 37:1167--1183.

Miglia, K., E. Mcarthur, W. Moore, H. Wang, J. Graham, and D. Freeman.
2005. Nine-year reciprocal transplant experiment in the gardens of the
basin and mountain big sagebrush (Artemisia tridentata: Asteraceae)
hybrid zone of Salt Creek Canyon: the importance of multiple-year
tracking of fitness Title. Biological Journal of the Linnean
Society:213--225.

Neilson, R., J. Lenihan, D. Bachelet, and R. Drapek. 2005. Climate
change implications for sagebrush ecosystems. Page 145 \emph{in} North
american wildlife and natural resources conference.

Pechanec, J., G. Pickford, and G. Stewart. 1937. Effects of the 1934
Drought on Native Vegetation of the Upper Snake River Plans, Idaho.
Ecology:490--505.

Perfors, T., J. Harte, and S. E. Alter. 2003. Enhanced growth of
sagebrush (Artemisia tridentata) in response to manipulated ecosystem
warming. Global Change Biology 9:736--742.

Petchey, O. L., M. Pontarp, T. M. Massie, S. K{é}fi, A. Ozgul, M.
Weilenmann, G. M. Palamara, F. Altermatt, B. Matthews, J. M. Levine, D.
Z. Childs, B. J. McGill, M. E. Schaepman, B. Schmid, P. Spaak, A. P.
Beckerman, F. Pennekamp, and I. S. Pearse. 2015. The ecological forecast
horizon, and examples of its uses and determinants. Ecology Letters
18:597--611.

Poore, R. E., C. A. Lamanna, J. J. Ebersole, and B. J. Enquist. 2009.
Controls on Radial Growth of Mountain Big Sagebrush and Implications for
Climate Change. Western North American Naturalist 69:556--562.

Queenborough, S. A., K. M. Burnet, W. J. Sutherland, A. R. Watkinson,
and R. P. Freckleton. 2011. From meso- to macroscale population
dynamics: A new density-structured approach. Methods in Ecology and
Evolution 2:289--302.

R Core Team. 2013. R: A language and environment for statistical
computing.

Rees, M., and S. P. Ellner. 2009. Integral projection models for
populations in temporally varying environments. Ecological Monographs
79:575--594.

Roberts, D. R., and A. Hamann. 2012. Predicting potential climate change
impacts with bioclimate envelope models: A palaeoecological perspective.
Global Ecology and Biogeography 21:121--133.

Ross, B. E., M. B. Hooten, J.-M. DeVink, and D. N. Koons. 2015. Combined
effects of climate, predation, and density dependence on Greater and
Lesser Scaup population dynamics. Ecological Applications 25:1606--1617.

Running, S. W., R. R. Nemani, F. A. Heinsch, M. Zhao, M. Reeves, and H.
Hashimoto. 2004. A Continuous Satellite-Derived Measure of Global
Terrestrial Primary Production. BioScience 54:547.

Rupp, D. E., J. T. Abatzoglou, K. C. Hegewisch, and P. W. Mote. 2013.
Evaluation of CMIP5 20th century climate simulations for the Pacific
Northwest US. Journal of Geophysical Research 118:1--23.

Salguero-G{ó}mez, R., O. R. Jones, C. R. Archer, Y. M. Buckley, J.
Che-Castaldo, H. Caswell, D. Hodgson, A. Scheuerlein, D. A. Conde, E.
Brinks, H. de Buhr, C. Farack, F. Gottschalk, A. Hartmann, A. Henning,
G. Hoppe, G. R{ö}mer, J. Runge, T. Ruoff, J. Wille, S. Zeh, R. Davison,
D. Vieregg, A. Baudisch, R. Altwegg, F. Colchero, M. Dong, H. de Kroon,
J.-D. Lebreton, C. J. E. Metcalf, M. M. Neel, I. M. Parker, T. Takada,
T. Valverde, L. A. V{é}lez-Espino, G. M. Wardle, M. Franco, and J. W.
Vaupel. 2015. The compadrePlant Matrix Database: an open online
repository for plant demography. Journal of Ecology 103:202--218.

Schlaepfer, D. R., W. K. Lauenroth, and J. B. Bradford. 2011.
Ecohydrological niche of sagebrush ecosystems. Ecohydrology:n/a--n/a.

Schlaepfer, D. R., W. K. Lauenroth, and J. B. Bradford. 2014a. Modeling
regeneration responses of big sagebrush (Artemisia tridentata) to
abiotic conditions. Ecological Modelling 286:66--77.

Schlaepfer, D., W. K. Lauenroth, and J. B. Bradford. 2012. Effects of
ecohydrological variables on current and future ranges, local
suitability patterns, and model accuracy in big sagebrush. Ecography
5:453--466.

Schlaepfer, D., W. K. Lauenroth, and J. B. Bradford. 2014b. Natural
Regeneration Processes in Big Sagebrush (Artemisia tridentata).
Rangeland Ecology \& Management 67:344--357.

>>>>

\documentclass[12pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\PassOptionsToPackage{usenames,dvipsnames}{color} % color is loaded by hyperref
\hypersetup{unicode=true,
            colorlinks=true,
            linkcolor=black,
            citecolor=Blue,
            urlcolor=black,
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

@@

\definecolor{blue}{rgb}{0,0,0}

@@

Forecasting the impacts of climate change on plant populations and
communities is a central challenge for ecology (Clark et al. 2001,
Petchey et al. 2015). Population models are ideally suited for meeting
such a challenge because they provide a way to link climate drivers
directly to population dynamics (Hare et al. 2010, Adler et al. 2012,
Ross et al. 2015, Shriver 2015). However, inference from population
models is typically limited to small spatial extents because the data
required is difficult to collect across broad species ranges. Almost
every study of plant population dynamics relies on demographic
observations recorded at the meter to sub-meter scale (see, e.g.,
Salguero-Gómez et al. 2015). Local-scale demographic data make building
population projection models an easy task (Ellner and Rees 2006, Rees
and Ellner 2009, Adler et al. 2012), but it is very difficult to
extrapolate small-scale studies to large spatial extents with any
certainty because the data likely only represent a small subset of
parameter space and environmental conditions (Freckleton et al. 2011,
Queenborough et al. 2011). The real challenge is not to simply make
population forecasts, but to do so at spatial scales relevant to policy
and management decisions (Queenborough et al. 2011).

@@

Large-scale, spatially-explicit population models based on RS data could
offer a valuable new way to investigate the effects of large-scale
environmental changes playing out at landscape and regional scales. Most
current assesments of how plant and animal populations will respond to
climate change rely on species distribution models (SDMs). SDMs rely on
static associations between contemporary climate and a species'
distribution or, more rarely, abundance to project future distribution
or abundance (Elith and Leathwick 2009) and they are easily applied at
landscape to continental scales (e.g., Maiorano et al. 2013, Clark et
al. 2014). However, the short-term and small-scale population dynamics
that actually drive the large-scale distributions of species are not
represented in most SDMs. Because SDMs typically rely on occurence data,
their projections of habitat suitability or probability of occurence
provide little information on the future states of populations in the
core of their range -- areas where a species exists now and is expected
to persist in the future (Ehrlén and Morris 2015). Furthermore, because
they lack short-term dynamics, SDMs usually cannot produce any estimate
of the rate at which local populations will increase or decrease in the
near-term and instead project a future equilibrium species distribution
that may or may not ever be reached. Direct validation of such
predictions is extremely rare (Roberts and Hamann 2012). Large-scale
dynamic population models could overcome these limitations. They would
produce spatially-explicit estimates of species abundance within the
species range (Ehrlén and Morris 2015), have the potential to model
expansion in abundance outside the range when coupled with dynamic
models of dispersal, and would provide testable predictions of how
populations should respond to short-term climate perturbations. These
short-term predictions also would give modelers the opportunity to
repeatedly validate and refine their models (Luo et al. 2011).

@@

\hypertarget{refs}{}
\hypertarget{ref-Adler2012}{}
Adler, P. B., H. J. Dalgleish, and S. P. Ellner. 2012. Forecasting plant
community impacts of climate variability and change: when do competitive
interactions matter? Journal of Ecology 100:478--487.

\hypertarget{ref-Apodaca2013}{}
Apodaca, L. F. 2013. Assessing Growth Response to Climate Controls in a
Great Basin Artemisia Tridentata Plant Community. PhD thesis, University
of Nevada Las Vegas.

\hypertarget{ref-Arnett2015}{}
Arnett, E. B., and T. Z. Riley. 2015. Science, policy, and the fate of
the greater sage-grouse. Frontiers in Ecology and the Environment
13:235.

\hypertarget{ref-Baldeck2014}{}
Baldeck, C. A., and G. P. Asner. 2014. Improving remote species
identification through efficient training data collection. Remote
Sensing 6:2682--2698.

\hypertarget{ref-Barry1996}{}
Barry, R. P., and J. M. V. Hoef. 1996. Blackbox Kriging: Spatial
Prediction without Specifying Variogram Models.

\hypertarget{ref-Bradford2006}{}
Bradford, J. B., and W. K. Lauenroth. 2006. Controls over invasion of
Bromus tectorum: The importance of climate, soil, disturbance, and seed
availability. Journal of Vegetation Science 17:693--704.

\hypertarget{ref-Bradley2010}{}
Bradley, B. A. 2010. Assessing ecosystem threats from global and
regional change: hierarchical modeling of risk to sagebrush ecosystems
from climate change, land use and invasive species in Nevada, USA.
Ecography 33:198--208.

\hypertarget{ref-Bradshaw2000}{}
Bradshaw, G. A., and J. G. Borchers. 2000. Uncertainty as information:
Narrowing the science-policy gap. Ecology and Society 4.

\hypertarget{ref-Clark2011}{}
Clark, J. S., D. M. Bell, M. H. Hersh, and L. Nichols. 2011. Climate
change vulnerability of forest biodiversity: Climate and competition
tracking of demographic rates. Global Change Biology 17:1834--1849.

\hypertarget{ref-Clark2001}{}
Clark, J. S., S. R. Carpenter, M. Barber, S. Collins, A. Dobson, J. A.
Foley, D. M. Lodge, M. Pascual, R. Pielke, W. Pizer, C. Pringle, W. V.
Reid, K. A. Rose, O. Sala, W. H. Schlesinger, D. H. Wall, and D. Wear.
2001. Ecological forecasts: an emerging imperative. Science
293:657--660.

\hypertarget{ref-Clark2014}{}
Clark, J. S., A. E. Gelfand, C. W. Woodall, and K. Zhu. 2014. More than
the sum of the parts: Forest climate response from joint species
distribution models. Ecological Applications 24:990--999.

\hypertarget{ref-Colgan2014}{}
Colgan, M. S., and G. P. Asner. 2014. Coexistence and environmental
filtering of species-specific biomass in an African savanna. Ecology
95:1579--1590.

\hypertarget{ref-Conn2015}{}
Conn, P. B., D. S. Johnson, J. M. V. Hoef, M. B. Hooten, J. M. London,
and P. L. Boveng. 2015. Using spatiotemporal statistical models to
estimate animal abundance and infer ecological dynamics from survey
counts. Ecological Monographs 85:235--252.

\hypertarget{ref-Dalgleish2011}{}
Dalgleish, H. J., D. N. Koons, M. B. Hooten, C. A. Moffet, and P. B.
Adler. 2011. Climate influences the demography of three dominant
sagebrush steppe plants. Ecology 92:75--85.

\hypertarget{ref-Debinski2010}{}
Debinski, D. M., H. Wickham, K. Kindscher, J. C. Caruthers, and M.
Germino. 2010. Montane meadow change during drought varies with
background hydrologic regime and plant functional group. Ecology
91:1672--81.

\hypertarget{ref-Ehrlen2015}{}
Ehrlén, J., and W. F. Morris. 2015. Predicting changes in the
distribution and abundance of species under environmental change.
Ecology Letters 18:303--314.

\hypertarget{ref-Elith2009}{}
Elith, J., and J. R. Leathwick. 2009. Species Distribution Models:
Ecological Explanation and Prediction Across Space and Time.

\hypertarget{ref-Ellner2006}{}
Ellner, S. P., and M. Rees. 2006. Integral projection models for species
with complex demography. The American naturalist 167:410--428.

\hypertarget{ref-Freckleton2011}{}
Freckleton, R. P., W. J. Sutherland, A. R. Watkinson, and S. A.
Queenborough. 2011. Density-structured models for plant population
dynamics. American Naturalist 177:1--17.

\hypertarget{ref-Gelman2009}{}
Gelman, A., and J. Hill. 2009. Data analysis using regression and
multilevel/hierarchical models. Cambridge University Press, Cambridge.

\hypertarget{ref-Gelman1992}{}
Gelman, A., and D. B. Rubin. 1992. Inference from Iterative Simulation
Using Multiple Sequences. Statistical Science 7:457--472.

\hypertarget{ref-Gerber2015}{}
Gerber, B. D., W. L. Kendall, M. B. Hooten, J. A. Dubovsky, and R. C.
Drewien. 2015. Optimal population prediction of sandhill crane
recruitment based on climate-mediated habitat limitations. Journal of
Animal Ecology 84:1299--1310.

\hypertarget{ref-Germino2014}{}
Germino, M. J., and K. Reinhardt. 2014. Desert shrub responses to
experimental modification of precipitation seasonality and soil depth:
Relationship to the two-layer hypothesis and ecohydrological niche.
Journal of Ecology 102:989--997.

\hypertarget{ref-Hare2010}{}
Hare, J. a, M. a Alexander, M. J. Fogarty, E. H. Williams, and J. D.
Scott. 2010. Forecasting the dynamics of a coastal fishery species using
a coupled climate--population model. Ecological applications : a
publication of the Ecological Society of America 20:452--464.

\hypertarget{ref-Harte2015}{}
Harte, J., S. R. Saleska, and C. Levy. 2015. Convergent ecosystem
responses to 23-year ambient and manipulated warming link advancing
snowmelt and shrub encroachment to transient and long-term climate-soil
carbon feedback. Global change biology 21:2349--56.

\hypertarget{ref-He2015}{}
He, K. S., B. A. Bradley, A. F. Cord, D. Rocchini, M.-N. Tuanmu, S.
Schmidtlein, W. Turner, M. Wegmann, and N. Pettorelli. 2015. Will remote
sensing shape the next generation of species distribution models? Remote
Sensing in Ecology and Conservation 1:4--18.

\hypertarget{ref-Higdon1998}{}
Higdon, D. 1998. A process-convolution approach to modelling
temperatures in the North Atlantic Ocean. Environmental and Ecological
Statistics 5:173--190.

\hypertarget{ref-Higdon2002}{}
Higdon, D. M. 2002. Space and space-time modeling using process
convolutions. Pages 37--56 \emph{in} C. Anderson, V. Barnett, P.
Chatwin, and A. El-Shaarawi, editors. Quantitative methods for current
environmental issues. Springer, London.

\hypertarget{ref-Hobbs2015}{}
Hobbs, N. T., and M. B. Hooten. 2015. Bayesian Models: A Statistical
Primer for Ecologists. Princeton University Press, Princeton.

\hypertarget{ref-Homer2012}{}
Homer, C. G., C. L. Aldridge, D. K. Meyer, and S. J. Schell. 2012.
Multi-scale remote sensing sagebrush characterization with regression
trees over Wyoming, USA: Laying a foundation for monitoring.
International Journal of Applied Earth Observation and Geoinformation
14:233--244.

\hypertarget{ref-Homer2015}{}
Homer, C. G., G. Xian, C. L. Aldridge, D. K. Meyer, T. R. Loveland, and
M. S. O'Donnell. 2015. Forecasting sagebrush ecosystem components and
greater sage-grouse habitat for 2050: Learning from past climate
patterns and Landsat imagery to predict the future. Ecological
Indicators 55:131--145.

\hypertarget{ref-Hooten2015}{}
Hooten, M. B., and N. T. Hobbs. 2015. A guide to Bayesian model
selection for ecologists. Ecological Monographs 85:3--28.

\hypertarget{ref-Hooten2007a}{}
Hooten, M. B., and C. K. Wikle. 2007. Shifts in the spatio-temporal
growth dynamics of shortleaf pine. Environmental and Ecological
Statistics 14:207--227.

\hypertarget{ref-Hooten2003}{}
Hooten, M. B., D. R. Larsen, and C. K. Wikle. 2003. Predicting the
spatial distribution of ground flora on large domains using a
hierarchical Bayesian model. Landscape Ecology 18:487--502.

\hypertarget{ref-Kuchler1964}{}
Kuchler, A. 1964. Potential Natural Vegetation of the Conterminous
United States. American Geographical Society, Special Publication No.
36.

\hypertarget{ref-Latimer2009}{}
Latimer, A. M., S. Banerjee, H. Sang, E. S. Mosher, and J. A. Silander.
2009. Hierarchical models facilitate spatial analysis of large data
sets: A case study on invasive plant species in the northeastern United
States. Ecology Letters 12:144--154.

\hypertarget{ref-Luo2011a}{}
Luo, Y., K. Ogle, C. Tucker, S. Fei, C. Gao, S. LaDeau, J. S. Clark, and
D. S. Schimel. 2011. Ecological forecasting and data assimilation in a
data-rich era. Ecological Applications 21:1429--1442.

\hypertarget{ref-Maiorano2013}{}
Maiorano, L., R. Cheddadi, N. E. Zimmermann, L. Pellissier, B.
Petitpierre, J. Pottier, H. Laborde, B. I. Hurdu, P. B. Pearman, A.
Psomas, J. S. Singarayer, O. Broennimann, P. Vittoz, A. Dubuis, M. E.
Edwards, H. A. Binney, and A. Guisan. 2013. Building the niche through
time: using 13,000 years of data to predict the effects of climate
change on three tree species in Europe. Global Ecology and Biogeography
22:302--317.

\hypertarget{ref-Merow2014}{}
Merow, C., A. M. Latimer, A. M. Wilson, S. M. McMahon, A. G. Rebelo, and
J. A. Silander. 2014. On using integral projection models to generate
demographically driven predictions of species' distributions:
development and validation using sparse data. Ecography 37:1167--1183.

\hypertarget{ref-Miglia2005}{}
Miglia, K., E. Mcarthur, W. Moore, H. Wang, J. Graham, and D. Freeman.
2005. Nine-year reciprocal transplant experiment in the gardens of the
basin and mountain big sagebrush (Artemisia tridentata: Asteraceae)
hybrid zone of Salt Creek Canyon: the importance of multiple-year
tracking of fitness Title. Biological Journal of the Linnean
Society:213--225.

\hypertarget{ref-Neilson2005}{}
Neilson, R., J. Lenihan, D. Bachelet, and R. Drapek. 2005. Climate
change implications for sagebrush ecosystems. Page 145 \emph{in} North
american wildlife and natural resources conference.

\hypertarget{ref-Pechanec1937}{}
Pechanec, J., G. Pickford, and G. Stewart. 1937. Effects of the 1934
Drought on Native Vegetation of the Upper Snake River Plans, Idaho.
Ecology:490--505.

\hypertarget{ref-Perfors2003}{}
Perfors, T., J. Harte, and S. E. Alter. 2003. Enhanced growth of
sagebrush (Artemisia tridentata) in response to manipulated ecosystem
warming. Global Change Biology 9:736--742.

\hypertarget{ref-Petchey2015}{}
Petchey, O. L., M. Pontarp, T. M. Massie, S. Kéfi, A. Ozgul, M.
Weilenmann, G. M. Palamara, F. Altermatt, B. Matthews, J. M. Levine, D.
Z. Childs, B. J. McGill, M. E. Schaepman, B. Schmid, P. Spaak, A. P.
Beckerman, F. Pennekamp, and I. S. Pearse. 2015. The ecological forecast
horizon, and examples of its uses and determinants. Ecology Letters
18:597--611.

\hypertarget{ref-Poore2009}{}
Poore, R. E., C. A. Lamanna, J. J. Ebersole, and B. J. Enquist. 2009.
Controls on Radial Growth of Mountain Big Sagebrush and Implications for
Climate Change. Western North American Naturalist 69:556--562.

\hypertarget{ref-Queenborough2011}{}
Queenborough, S. A., K. M. Burnet, W. J. Sutherland, A. R. Watkinson,
and R. P. Freckleton. 2011. From meso- to macroscale population
dynamics: A new density-structured approach. Methods in Ecology and
Evolution 2:289--302.

\hypertarget{ref-R2013}{}
R Core Team. 2013. R: A language and environment for statistical
computing.

\hypertarget{ref-Rees2009}{}
Rees, M., and S. P. Ellner. 2009. Integral projection models for
populations in temporally varying environments. Ecological Monographs
79:575--594.

\hypertarget{ref-Roberts2012}{}
Roberts, D. R., and A. Hamann. 2012. Predicting potential climate change
impacts with bioclimate envelope models: A palaeoecological perspective.
Global Ecology and Biogeography 21:121--133.

\hypertarget{ref-Ross2015}{}
Ross, B. E., M. B. Hooten, J.-M. DeVink, and D. N. Koons. 2015. Combined
effects of climate, predation, and density dependence on Greater and
Lesser Scaup population dynamics. Ecological Applications 25:1606--1617.

\hypertarget{ref-Running2004}{}
Running, S. W., R. R. Nemani, F. A. Heinsch, M. Zhao, M. Reeves, and H.
Hashimoto. 2004. A Continuous Satellite-Derived Measure of Global
Terrestrial Primary Production. BioScience 54:547.
